Millions of Type Errors

### 2023-09-07 programming23/paper.tex
counting specific type errors

\usepackage{alltt}
\usepackage[backend=biber]{biblatex}
\usepackage{amssymb}
\usepackage{calc}
\usepackage{listings}
\usepackage{mathpartir}
\usepackage{pifont}
\usepackage{siunitx}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{wasysym}
\usepackage{xcolor}
\usetikzlibrary{shapes.geometric}

@@

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\stddev}[1]{[#1\textsc{k}]}

@@

\begin{itemize}
\item
  No personal data (PII)
  \subitem
    Fairly obvious.
    But, we also can't collect directory names, filenames, or code snapshots
    because these all might have PII throughout.
\item
  No data about public events.
  \subitem
    Contributions to the Roblox Marketplace, for instance, are a public event.
    Although it would be interesting to know the number of type errors
    in published experiences, this data might be used to identify creators
    who publish often --- or on a distinct schedule.
\item
  Low cost
  \subitem
    Building, sending, and storing the data must all run quickly.
    FILL alan: what was our hard cap on Kibana use?
\end{itemize}

@@

  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll} \\
      \begin{tabular}[t]{l@{~~}r@{~}l}
        313,509 & \mnocheck{}   & [\pct{90.19}] \\
         32,902 & \mnonstrict{} & [\pct{ 9.47}] \\
            545 & \mstrict{}    & [\pct{ 0.16}] \\
            642 & mixed-mode    & [\pct{ 0.18}]
      \end{tabular}
      \begin{tabular}[t]{l@{~~}ll}
        \zerowidth{Among the mixed-mode sessions:} \\
        & 341 & contain a mode upgrade \\
        & 320 & contain a mode downgrade \\
        & 512 & contain modules with different modes
      \end{tabular}
    \end{tabular}
    \caption{Total sessions (347,598), whether/how they switch analysis mode}
    \label{f:total-sessions}
  \end{subfigure}

@@

%% TODO power-law: event count, timespan

FILL te in edit range, 
 14,917 non-stx
 16,007 syntax in edit ranges.
 Most are typos!

%% results 1 section?!

%% TODO counting stats on type errors
%% - specific errors
%% - typos vs rest
%% - too complex
%% - mod switch vs other

\begin{figure}[t]
  %% TODO stacked bars?
  %% TODO align old;curr
  %% TODO sep23 remove this??!
  \begin{tabular}{lrrrrr}
    Mode             & \# Rows & E         & Any Error & In Module & In range \\
      & & & & (old ; curr) &  (old ; curr) \\\midrule
    \mnocheck{}   &   4,471 & \code{te} &    5.77\% &  0.20\% ;  2.21\% &  0.16\% ;  0.63\% \\
                  &         & \code{fs} &   59.18\% & 32.81\% ; 49.16\% &  9.17\% ; 13.13\% \\
    \mnonstrict{} &     506 & \code{te} &   25.89\% & 12.06\% ; 18.18\% &  6.32\% ;  7.91\% \\
                  &         & \code{fs} &   62.45\% & 37.94\% ; 53.75\% & 16.60\% ; 20.75\% \\
    \mstrict{}    &      24 & \code{te} &   50.00\% & 29.17\% ; 33.33\% &     0\% ;     0\% \\
                  &         & \code{fs} &   62.50\% & 50.00\% ; 58.33\% &  4.17\% ;  8.33\% 
  \end{tabular}

  \bigskip

  Below, edit range only.
  Each record with an edit range error adds +1 to at least 1 row.

@@

\Cref{f:dataset-overview}:
cannot filter syntax errors from the total and per-module error counts (\cref{s:threats});
multi-mode project = found a module switch that coincides with a mode switch, which suggests that two modules have different modes;
upgrade = found a up-mode switch that does not match a module switch (quite possible to miss these);
downgrade = found a down-mode switch that does not match a module switch

@@

>>>>

\usepackage{alltt}
\usepackage[backend=biber]{biblatex}
\usepackage{amssymb}
\usepackage{calc}
\usepackage{colortbl}
\usepackage{listings}
\usepackage{mathpartir}
\usepackage{pifont}
\usepackage{siunitx}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{wasysym}
\usepackage{xcolor}
\usetikzlibrary{shapes.geometric}

@@

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\stddev}[1]{[#1\textsc{k}]}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}

@@

\begin{itemize}
\item
  No personal data (PII)
  \subitem
    Fairly obvious.
    But, we also can't collect directory names, filenames, or code snapshots
    because these all might have PII throughout.
\item
  No data about public events.
  \subitem
    Contributions to the Roblox Marketplace, for instance, are a public event.
    Although it would be interesting to know the number of type errors
    in published experiences, this data might be used to identify creators
    who publish often --- or on a distinct schedule.
\item
  Low cost
  \subitem
    Building, sending, and storing the data must all run quickly.
    FILL what was our hard cap on Kibana use?
\end{itemize}

@@

  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll} \\
      \begin{tabular}[t]{l@{~~}r@{~}l}
        313,509 & \mnocheck{}   & [\pct{90.19}] \\
         32,902 & \mnonstrict{} & [\pct{ 9.47}] \\
            545 & \mstrict{}    & [\pct{ 0.16}] \\
            642 & mixed-mode    & [\pct{ 0.18}]
      \end{tabular}
      \begin{tabular}[t]{l@{~~}ll}
        \zerowidth{Among the mixed-mode sessions:} \\
        & 341 & contain a mode upgrade \\
        & 320 & contain a mode downgrade \\
        & 512 & contain modules with different modes
      \end{tabular}
    \end{tabular}
    \caption{Total sessions (347,598), whether and how they switch analysis mode}
    \label{f:total-sessions}
  \end{subfigure}

@@

\Cref{f:dataset-overview}

FILL nocheck nonstrict strict, order-of-magnitude

FILL type error counts
Mostly syntax errors though!
Estimate \pct{40} are substantial type errors rather than syntax
errors based on the edit range type errors~(\cref{s:type-error-count})
but unfortunately we cannot get an exact number for the overall
or per-module type error counts.

FILL single-mode sessions, mostly

FILL details on upgrades, downgrades, then ignore these for rest of paper.
multi-mode project = found a module switch that coincides with a mode switch, which suggests that two modules have different modes;
upgrade = found a up-mode switch that does not match a module switch (quite possible to miss these);
downgrade = found a down-mode switch that does not match a module switch

FILL most sessions fairly long in time (order of minutes),
short in event count (dozens).

\section{Type Errors in Edit Range}
\label{s:type-error-count}

\begin{table}[t]
  %% Looks the same when restricted to modules
  \begin{tabular}{l@{~~}l}
    \begin{tabular}[t]{lr}
      \code{SyntaxError} & 60.79\% \\
      \code{UnknownSymbol} & 27.34\% \\
      \code{UnknownProperty} & 4.08\% \\
      \code{TypeMismatch} & 1.92\% \\
      \code{UnknownRequire} & 1.40\% \\
      \code{CountMismatch} & 1.08\% \\
      \code{OptionalValueAccess} & 0.99\% \\
      \code{GenericError} & 0.93\% \\
      \code{CannotExtendTable} & 0.21\% \\
      \code{UnknownPropButFoundLikeProp} & 0.20\% \\
      \code{FunctionExitsWithoutReturning} & 0.16\% \\
      \code{ExtraInformation} & 0.14\% \\
      \code{CannotInferBinaryOperation} & 0.14\% \\
      \code{MissingProperties} & 0.13\% \\
      \code{ModuleHasCyclicDependency} & 0.10\% \\
      \code{IllegalRequire} & 0.09\% \\
      \code{NotATable} & 0.09\% \\
      \code{MissingUnionProperty} & 0.08\%
    \end{tabular} &
    \begin{tabular}[t]{lr}
      \code{CannotCallNonFunction} & 0.07\% \\
      \code{FunctionDoesNotTakeSelf} & 0.03\% \\
      \code{IncorrectGenericParameterCount} & 0.01\% \\
      \code{OnlyTablesCanHaveMethods} & 0.00\% \\
      \code{TypesAreUnrelated} & 0.00\% \\
      \code{DuplicateTypeDefinition} & 0.00\% \\
      \code{OccursCheckFailed} & 0.00\% \\
      \code{TypePackMismatch} & 0.00\% \\
      \code{FunctionRequiresSelf} & 0.00\% \\
      \code{DuplicateGenericParameter} & 0.00\% \\
      \code{InternalError} & 0.00\% \\
      \code{CodeTooComplex} & 0.00\% \\
      \code{DeprecatedApiUsed} & 0.00\% \\
      \code{NormalizationTooComplex} & 0.00\% \\
      \code{SwappedGenericTypeParameter} & 0.00\% \\
      \code{DynamicPropertyLookupOnClassesUnsafe} & 0.00\% \\
      \code{UnificationTooComplex} & 0.00\% \\
    \end{tabular}
  \end{tabular}
  \caption{Type error popularity}
  \label{t:type-error-count}
\end{table}

FILL popularity over all old and new in edit range.
Syntax error by far the most common.
UnknownSymbol is likely a syntax error --- so \pct{90} syntax error!

Fourteen errors never showed up.
Nothing for \code{CodeTooComplex} or \code{UnificationTooComplex} is great news.
But what about the others? FILL

=== BEN START HERE

TODO get counts for each mode. Are there fewer stx errors in strict mode??!

TODO which type errors at end of session?

\begin{figure}[t]
  Below, edit range only.
  Each record with an edit range error adds +1 to at least 1 row.

@@

@@

\section{Analysis}

\begin{figure}[thb]
  %% TODO stacked bars?
  %% TODO align old;curr
  %% TODO sep'23 remove this??!
  \begin{tabular}{lrrrrr}
    Mode             & \# Rows & E         & Any Error & In Module & In range \\
      & & & & (old ; curr) &  (old ; curr) \\\midrule
    \mnocheck{}   &   4,471 & \code{te} &    5.77\% &  0.20\% ;  2.21\% &  0.16\% ;  0.63\% \\
                  &         & \code{fs} &   59.18\% & 32.81\% ; 49.16\% &  9.17\% ; 13.13\% \\
    \mnonstrict{} &     506 & \code{te} &   25.89\% & 12.06\% ; 18.18\% &  6.32\% ;  7.91\% \\
                  &         & \code{fs} &   62.45\% & 37.94\% ; 53.75\% & 16.60\% ; 20.75\% \\
    \mstrict{}    &      24 & \code{te} &   50.00\% & 29.17\% ; 33.33\% &     0\% ;     0\% \\
                  &         & \code{fs} &   62.50\% & 50.00\% ; 58.33\% &  4.17\% ;  8.33\% 
  \end{tabular}

  \caption{TODO}

\end{figure}
